---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# hpa.spatial

<!-- badges: start -->
<!-- badges: end -->

The goal of hpa.spatial is to make relevant shape files and data easily available and include helpful functions for analysis of spatial data.

## Notes on other packages

Shape files are available within [`{absmapsdata}`](https://github.com/wfmackey/absmapsdata) and these can be (lazy) loaded using [`{strayr}`](https://github.com/runapp-aus/strayr).





## Installation

You can install the development version of hpa.spatial from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("healthpolicyanalysis/hpa.spatial")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(hpa.spatial)
## basic example code
```

