---
title: "Advanced mapping with correspondence example - ACPR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced-mapping-with-correspondence-example-ACPR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(hpa.spatial)
library(sf)
library(dplyr)
library(ggplot2)
```

# Current problem

- The year is 2024. 
- We want to map count data from SA2s (2021 edition) to Aged Care Planning Regions (ACPRs). 
- There has been population growth in Australia since the 2021 census population (which the `map_with_correspondence()` uses by default for creating correspondence tables).


### The plan

Use `map_with_correspondence()` but create a new `sf (POINT)` object that incorporates the estimated residential population projections (ERP) from 2023 at the SA2 level. Since we don't have mesh block population data from the ERP, this won't account for within-SA2-differences in population growth from 2021 to 2023, but it will account for differences in between-SA2-differences in population growth.



```{r}
sa2_erp23 # use ERP data at SA2 level from ABS <https://www.abs.gov.au/statistics/people/population/regional-population/latest-release#data-downloads>

mb21_pop <- get_mb21_pop()

# create adjusted MB21 population data for creating correspondence table using ERP data at SA2 level
adj_mb21_pop <- mb21_pop |>
  as_tibble() |>
  select(MB_CODE21, sa2_code_2021 = SA2_CODE21, Person) |>
  # create within-sa2 pop ratios for each mesh block
  mutate(within_sa2_pp = Person / sum(Person), .by = sa2_code_2021) |>
  # join the ERP (2023) at SA2 level
  inner_join(sa2_erp23, by = "sa2_code_2021") |>
  # apply SA2 level ERP to within-sa2 pop ratios to get adjusted pop at mesh block level
  mutate(pop23 = erp * within_sa2_pp) |>
  select(MB_CODE21, Person = pop23) |>
  # join the POINT geometry back
  (\(d) {
    mb21_pop |>
      select(MB_CODE21, sa2_code_2021 = SA2_CODE21) |>
      inner_join(d, by = "MB_CODE21")
  })()

adj_mb21_pop
```


```{r}
# create SA2 count data for mapping
d_sa2 <- get_polygon("sa22021") |>
  as_tibble() |>
  select(sa2_code_2021) |>
  filter(sa2_code_2021 %in% adj_mb21_pop$sa2_code_2021)

d_sa2$measure <- rpois(n = nrow(d_sa2), lambda = 1000)


# map (and aggregate) the data to ACPR
mapped_measures <- map_data_with_correspondence(
  .data = d_sa2,
  codes = sa2_code_2021,
  values = measure,
  from_geo = get_polygon("sa22021"),
  to_geo = get_polygon("ACPR"),
  mb_geo = adj_mb21_pop,
  export_fname = "sa22021-to-acpr",
  value_type = "aggs"
)

# map (and aggregate) the ERP to ACPR so that a rate can be calculated
mapped_pop <- map_data_with_correspondence(
  .data = sa2_erp23,
  codes = sa2_code_2021,
  values = erp,
  from_geo = get_polygon("sa22021"),
  to_geo = get_polygon("ACPR"),
  mb_geo = adj_mb21_pop,
  export_fname = "sa22021-to-acpr",
  value_type = "aggs"
)

sa2_rate_poly <- get_polygon("sa22021", crs = 7844) |>
  inner_join(d_sa2, by = "sa2_code_2021") |>
  left_join(sa2_erp23, by = "sa2_code_2021") |>
  mutate(rate = measure / erp)

acpr_rate_poly <- get_polygon("ACPR", crs = 7844) |>
  inner_join(mapped_measures, by = "acpr_code") |>
  left_join(mapped_pop, by = "acpr_code") |>
  mutate(rate = measure / erp)

bind_rows(
  select(sa2_rate_poly, rate, erp) |> mutate(geo = "SA2 (2021)"),
  select(acpr_rate_poly, rate, erp) |> mutate(geo = "ACPR (2018)")
) |>
  filter(erp > 0) |>
  ggplot() +
  geom_sf(aes(fill = rate), col = NA) +
  scale_fill_continuous(trans = "log10", labels = scales::label_comma()) +
  facet_wrap(~geo, ncol = 1) +
  theme_bw() +
  labs(fill = "Rate")
```

